\chapter{Introduction}

SeedBot is an automated system designed for agricultural sowing. It uses a network of wireless sensors to monitor soil conditions and control the seed distribution mechanism, improving efficiency and reducing manual labor. Using a pre-trained machine learning model, SeedBot can determine the most suitable crop for a specific plot of land.

\subsection{Use Case}
The agricultural sector plays a crucial role in the industry, specifically when considering large-scale food production, which requires high efficiency, a strict quality check and optimized comunication among the different phases of the production process. However, traditional farming methods are labor-intensive and often inefficient. 
In the contest of smart industries, the use of IoT devices and machine learning algorithms can significantly improve agricultural processes.\\


SeedBot aims to address these challenges by automating the sowing process, optimizing seed distribution based on real-time soil data, and reducing the need for manual intervention. This allows the sowing to happen in in optimal conditions, minimazing errors and the waste of resources.\\

SeedBot receives real-time data from various sensors, including nitrogen, phosphorus, and potassium (NPK) levels, soil moisture, pH, and air temperature. This data is used to determine the best seed type for the soil conditions and control the seed distribution mechanism accordingly. The position and the results of the sowing are then sent to a central server for monitoring and analysis, which allows farmers to correct anomalies and optimize the sowing process.\\

This means that SeedBot allows to reduce operational costs and production time, helping to make the farms more competitive on the global market.