\chapter{Conclusion}

\subsection{Future Implementation}

In the future, SeedBot could be improved by including:

\begin{itemize}
    \item \textbf{Advanced machine learning algorithms:} To further optimize seed distribution and adapt to various weather and soil conditions, enabling predictive analytics for sowing times and crop yields.
    \item \textbf{Advanced remote user Interface:} Using platforms like Grafana to monitor sensor data in real-time and provide a more interactive user interface, allowing users to set thresholds for various parameters (e.g., soil moisture, pH levels) and receive automated alerts.
    \item \textbf{Integration with irrigation and nutrition systems:} To provide a complete crop management system that includes not only sowing but also irrigation and plant nutrition, with intelligent scheduling based on environmental data.
    \item \textbf{Energy efficiency and sustainability improvements:} Introducing low-power modes and energy-efficient protocols to extend the battery life of the devices and explore the use of renewable energy sources like solar panels to power the system in the field.
    \item \textbf{Blockchain for data integrity:} Utilizing blockchain technology to ensure the integrity of data collected from the sensors, providing transparent and tamper-proof records of farming activities, useful for certifications and traceability.
\end{itemize}

\subsection{Conclusion}

This project has demonstrated that handling advanced technologies, such as IoT devices and machine learning, can greatly improve the agricultural sector. The power of nRF52840 devices and the CoAP protocol helps to utilize the resources in an efficient and precise way.
Automating key processes like soil monitoring and seed sowing, means that the need for manual intervention reduces significantly. This approach has the potential to revolutionize farming practices, making them more sustainable and productive. The successful implementation of this system paves the way for further innovations in smart agriculture, ultimately contributing to a more efficient and data-driven approach to farming.